---
title: "Project 1-API-Vignette"
author: "Chennade Brown"
date: "9/26/2021"
output:
  github_document
    code_folding: show
    default_print: tibble
--

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Interacting with API's Vignette: Example with the Covid19 API.
This vignette will go through the steps of reading and summarizing data from an API. We will utilize the Covid19 API.

## Packages Required 
The following packages were used to utilize the functions needed for interacting with the API and manipulating the retrieved data:  
* `httr`: Used to retrieve data from APIs.   
* `jsonlite`: Functions used to manipulate JSON data.  
* `tidyverse`: Functions used to get manipulate and reshape data.  

```{r, echo = TRUE, results = 'hide', message = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
```

### Accessing the Data
First we will access data from the API regarding confirmed Covid 19 Cases for Norway and Switzerland during the time frame from March 1, 2020 - April 1, 2020 for comparison.  This will be achieved using the GET function from the `httr` package to return information from the API.
```{r, echo = TRUE}
getCountry <- GET("https://api.covid19api.com/country/Switzerland/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

getCountry2 <- GET("https://api.covid19api.com/country/Norway/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

getCountry
getCountry2
```

We will now review the contents of the returned data. The content function with the text parameter converts the raw data to JSON.
```{r, echo = TRUE}
getCountryText <- content(getCountry, "text")
getCountryText2 <- content(getCountry2, "text")
getCountryText
getCountryText2
```

We will now covert the raw data into JSON format which will allow us to PARSE the data and output a data frame using the jsonlite package that was previously installed.
```{r, echo = TRUE}
getCountryJs <- fromJSON(getCountryText, flatten = TRUE)
getCountryJs2 <- fromJSON(getCountryText2, flatten = TRUE)
```

We will convert the parsed JSON data for Norway & Switzerland into a tibble for our analysis.
```{r, echo = TRUE}
getCountryJsTb <- as_tibble(getCountryJs)
getCountryJsTb
```

```{r, echo = TRUE}
getCountryJs2Tb <- as_tibble(getCountryJs2)
getCountryJs2Tb
```

We will now combine the data for Norway and Switzerland into one data frame using the rbind function and remove the empty columns Province, City, & CityCode.
```{r, echo = TRUE}
NorSwiz <- function(df1, df2) {rbind(df1, df2)}
Combo <- NorSwiz(df1 = getCountryJsTb, df2 = getCountryJs2Tb)
Combo$Province <- NULL
Combo$City <- NULL
Combo$CityCode <- NULL
Combo
```