
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Introduction
This vignette will go through the steps of reading and summarizing data from an API. We will utilize the Covid19 API.

# Packages Required 
The following packages were used to utilize the functions needed for interacting with the API and manipulating the retrieved data:  
* `httr`: Used to retrieve data from APIs.   
* `jsonlite`: Functions used to manipulate JSON data.  
* `tidyverse`: Functions used to manipulate and reshape data.  
* `dplyr`: Function that contains a set of tools for manipulating data in R.

```{r, echo = TRUE, results = 'hide', message = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
```

# Accessing the Data
First we will access data from the API regarding confirmed Covid 19 Cases for Norway and Switzerland during the time frame from March 1, 2020 - April 1, 2020 for comparison.  This will be achieved using the GET function from the `httr` package to return information from the API.
```{r, echo = TRUE}
getCountry <- GET("https://api.covid19api.com/country/Switzerland/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

getCountry2 <- GET("https://api.covid19api.com/country/Norway/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

```

We will now review the contents of the returned data. The content function with the text parameter converts the raw data to JSON.
```{r, echo = TRUE}
getCountryText <- content(getCountry, "text")
getCountryText2 <- content(getCountry2, "text")
```

We will now covert the raw data into JSON format which will allow us to PARSE the data and output a data frame using the jsonlite package that was previously installed.
```{r, echo = TRUE}
getCountryJs <- fromJSON(getCountryText, flatten = TRUE)
getCountryJs2 <- fromJSON(getCountryText2, flatten = TRUE)
```

We will convert the parsed JSON data for Norway & Switzerland into a tibble for our analysis.
```{r, echo = TRUE}
getCountryJsTb <- as_tibble(getCountryJs)
```

```{r, echo = TRUE}
getCountryJs2Tb <- as_tibble(getCountryJs2)
```

We will now combine the data for Norway and Switzerland into one data frame using the rbind function and remove the empty columns Province, City, & CityCode.
```{r, echo = TRUE}
NorSwiz <- function(df1, df2) {rbind(df1, df2)}
Combo <- NorSwiz(df1 = getCountryJsTb, df2 = getCountryJs2Tb)
Combo$Province <- NULL
Combo$City <- NULL
Combo$CityCode <- NULL
```

The following functions will allow the user to query the Norway and Switzerland API by entering the column names or all.
```{r, echo = TRUE, message = FALSE, results = 'hide'}
ConfirmedSwiss <- function(type = "all"){
  OutputDatSwiss <- GET("https://api.covid19api.com/country/Switzerland/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")
    DataSwiss <- fromJSON(rawToChar(OutputDatSwiss$content))
  if(type!="all"){DataSwiss <- DataSwiss %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date, Status)}
  return(DataSwiss)}
ConfirmedSwiss("Country")
```
```{r, echo = TRUE, message = FALSE, results = 'hide'}
ConfirmedNor <- function(type = "all"){
  OutputDatNor <- GET("https://api.covid19api.com/country/Norway/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")
    DataNor <- fromJSON(rawToChar(OutputDatNor$content))
  if(type!="all"){DataNor <- DataNor %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date, Status)}
  return(DataNor)}
ConfirmedNor("CountryCode")
```

The following functions return Day One Data for Africa & Mexico for the first recorded case. We will go through the same steps outlined above to retrieve and parse the data from the API.
```{r, echo = TRUE, message = FALSE}
DayOneDataSa <- GET(
  url = "https://api.covid19api.com/dayone/country/south-africa/status/confirmed")
DayOneDataMex <- GET(
  url = "https://api.covid19api.com/dayone/country/mexico/status/confirmed")

DayOneDataText <- content(DayOneDataSa, "text")
DayOneDataTextMex <- content(DayOneDataMex, "text")

DayOneDataJson <- fromJSON(DayOneDataText, flatten = TRUE)
DayOneDataJsonMex <- fromJSON(DayOneDataTextMex, flatten = TRUE)

DayOneDfSa <- as_tibble(DayOneDataJson)
DayOneDfMex <- as_tibble(DayOneDataJsonMex)

```
We will combine the data sets for the analysis.
```{r, echo = TRUE}
SaMex <- function(df1, df2) {rbind(df1, df2)}
Day1 <- SaMex(df1 = DayOneDfSa, df2 = DayOneDfMex)

Day1$Province <- NULL
Day1$City <- NULL
Day1$CityCode <- NULL
```

These functions will allow the user to query the Day One Mexico and Day One South Africa APIs based on the columns entered by the user.  The Day1 data set will be used in the analysis.
```{r, echo = TRUE, message = FALSE, results = 'hide'}
DayOneSa <- function(type = "all"){
  OutputDat <- GET("https://api.covid19api.com/dayone/country/south-africa/status/confirmed")
    DataSa <- fromJSON(rawToChar(OutputDat$content))
  if(type!="all"){DataSa <- DataSa %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date)}
  return(DataSa)}
DayOneSa("CountryCode")
```
```{r, echo = TRUE, message = FALSE, results = 'hide'}
DayOneMex <- function(type = "all"){
  OutputDatMex <- GET("https://api.covid19api.com/dayone/country/mexico/status/confirmed")
    DataMex <- fromJSON(rawToChar(OutputDatMex$content))
  if(type!="all"){DataMex <- DataMex %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date)}
  return(DataMex)}
DayOneMex("CountryCode")
```

We will now look at summary data which contains a summary of new and total cases per country which is updated daily utilizing the steps outlined above to retrieve and parse data from the API.
```{r, echo = TRUE, message = FALSE}
resp2 <- GET("https://api.covid19api.com/summary")

resp2Text <- content(resp2, "text")
resp2Json <- fromJSON(resp2Text, flatten = TRUE)

# Save data frame as resp1Df and remove the ID column.
resp2Df <- as.data.frame(resp2Json$Countries)

resp2Df$ID <- NULL
```

```{r, echo = TRUE, message = FALSE, results = 'hide'}
# This is a function allows the user to return NewDeaths, TotalDeaths,& NewRecovered, or all of the data based on inputs from the Summary API. We will use the above returned data set resp2Df for the analysis.

Summary <- function(type = "all"){
  OutputAPI <- GET("https://api.covid19api.com/summary")
    data <- fromJSON(rawToChar(OutputAPI$content))
    data2 <- data$Countries
  if(type!="all"){data2 <- data2 %>% select(type, Country, CountryCode,  NewDeaths, TotalDeaths, NewRecovered)}
  return(data2)}
Summary("Country")
```

# Data Analysis
We will create a contingency table that shows the occurrences of confirmed cases  between March - April 2020 for Switzerland and Norway.  Each country had a confirmed case during this time frame.
```{r, echo = TRUE}
tbl <- table(Combo$Country, Combo$Status)
tbl
```

Below is a box plot representing confirmed cases in Norway and Switzerland between March - April 2020.  The output shows that Switzerland had a higher number of confirmed cases.
```{r, echo = TRUE}
ggplot(Combo, aes(x = Cases, y = Country)) +
geom_boxplot() + geom_jitter(aes(color = Status)) + ggtitle("Boxplot for Confirmed Cases")

```
The following code calculates numerical summaries for daily cases confirmed for the two countries.  The mean case count per day for Norway was 1,759 cases and the mean case count per day for Switzerland was 5,402 cases. 
```{r, echo = TRUE}
Combo %>%
  group_by(Country) %>%
  summarize(Avg = mean(Cases), Sd = sd(Cases), Median = median(Cases), IQR =    IQR(Cases))

```
The following contingency table reports the number of confirmed case statuses  for south Africa and Mexico. 
```{r, echo = TRUE}
tbl2 <- table(Day1$Status, Day1$Country)
tbl2
```

The following bar graph reports the number of confirmed case statuses for South Africa and Mexico since day one which were similar.  
```{r, echo = TRUE}
ggplot(Day1, aes(x = Country)) + geom_bar(aes(fill = Status), position = "dodge") + xlab("Country") + scale_fill_discrete(name = "") + ggtitle("Confirmed Case Statuses")

```

The following code calculates numerical summaries for daily cases confirmed since Day 1 for the two countries.  The mean case count for Mexico was 1,418,233 cases and the mean case count for South Africa was 1,115,631 cases.
```{r, echo = TRUE}
Day1 %>%
  group_by(Country) %>%
  summarize(Avg = mean(Cases), Sd = sd(Cases), Median = median(Cases), IQR =    IQR(Cases))
```

The following code creates a new variable (NewRatio = TotalDeaths/Total Confirmed) that represents the ratio of total deaths to total confirmed cases and appends to the Resp2Df which is based on total cases for Countries that is updated daily.
```{r, echo = TRUE}
resp2Df <- resp2Df %>% mutate(Ratio = TotalDeaths/TotalConfirmed)
```

The following code returns numerical summaries for total deaths among all countries reported in this Summary API.  The average total deaths was 24,941.
```{r, echo = TRUE}
resp2Df %>%
    summarize(Avg = mean(TotalDeaths), Max = max(TotalDeaths), Sd = sd(TotalDeaths), Min = min(TotalDeaths))

```

The following code returns a scatter plot for the summary data set.  The plot shows there is minimal data on Newly Confirmed Cases and New Deaths; however, the graph shows a positive correlation between newly confirmed cases and new deaths.
```{r, echo = TRUE}
g <- ggplot(resp2Df, aes(x = NewConfirmed, y = NewDeaths))+ labs(y="New Deaths", x = "New Confirmed")
g + geom_point() + ggtitle("Confirmed Cases vs Deaths")
```

