---
title: "Covid19 API Vignette"
author: "Chennade Brown"
date: "10/03/2021"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Introduction
This vignette will go through the steps of reading and summarizing data from an API. We will utilize the Covid19 API.

# Packages Required 
The following packages were used to utilize the functions needed for interacting with the API and manipulating the retrieved data:  
* `httr`: Used to retrieve data from APIs.   
* `jsonlite`: Functions used to manipulate JSON data.  
* `tidyverse`: Functions used to manipulate and reshape data.  
* `dplyr`: Function that contains a set of tools for manipulating data in R.

```{r, echo = TRUE, results = 'hide', message = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
```

# Accessing the Data
First we will access data from the API regarding confirmed Covid 19 Cases for Norway and Switzerland during the time frame from March 1, 2020 - April 1, 2020 for comparison.  This will be achieved using the GET function from the `httr` package to return information from the API.
```{r, echo = TRUE}
getCountry <- GET("https://api.covid19api.com/country/Switzerland/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

getCountry2 <- GET("https://api.covid19api.com/country/Norway/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")

```

We will now review the contents of the returned data. The content function with the text parameter converts the raw data to JSON.
```{r, echo = TRUE}
getCountryText <- content(getCountry, "text")
getCountryText2 <- content(getCountry2, "text")
```

We will now covert the raw data into JSON format which will allow us to PARSE the data and output a data frame using the jsonlite package.
```{r, echo = TRUE}
getCountryJs <- fromJSON(getCountryText, flatten = TRUE)
getCountryJs2 <- fromJSON(getCountryText2, flatten = TRUE)
```

We will convert the parsed JSON data for Norway & Switzerland into a tibble for our analysis.
```{r, echo = TRUE}
getCountryJsTb <- as_tibble(getCountryJs)
```

```{r, echo = TRUE}
getCountryJs2Tb <- as_tibble(getCountryJs2)
```

We will now combine the data for Norway and Switzerland into one data frame using the rbind function and remove the empty columns Province, City, & CityCode.
```{r, echo = TRUE}
NorSwiz <- function(df1, df2) {rbind(df1, df2)}
Combo <- NorSwiz(df1 = getCountryJsTb, df2 = getCountryJs2Tb)
Combo$Province <- NULL
Combo$City <- NULL
Combo$CityCode <- NULL
Combo
```

The following functions will allow the user to query the Norway and Switzerland API by entering the column names or all.  To subset by rows the user can enter observations for the filter function.
```{r, echo = TRUE, message = FALSE, results = 'hide'}
ConfirmedSwiss <- function(type = "all"){
  OutputDatSwiss <- GET("https://api.covid19api.com/country/Switzerland/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")
    DataSwiss <- fromJSON(rawToChar(OutputDatSwiss$content))
  if(type!="all"){DataSwiss <- DataSwiss %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date, Status) %>% filter(Cases == 27)}
  return(DataSwiss)}
ConfirmedSwiss("Country")
```
```{r, echo = TRUE, message = FALSE, results = 'hide'}
ConfirmedNor <- function(type = "all"){
  OutputDatNor <- GET("https://api.covid19api.com/country/Norway/status/confirmed?from=2020-03-01T00:00:00Z&to=2020-04-01T00:00:00Z")
    DataNor <- fromJSON(rawToChar(OutputDatNor$content))
  if(type!="all"){DataNor <- DataNor %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date, Status) %>% filter(Cases == 19)}
  return(DataNor)}
ConfirmedNor("CountryCode")
```

The following functions return Day One data for Africa & Mexico for the first recorded case. We will go through the same steps outlined above to retrieve and parse the data from the API.
```{r, echo = TRUE, message = FALSE}
DayOneDataSa <- GET(
  url = "https://api.covid19api.com/dayone/country/south-africa/status/confirmed")
DayOneDataMex <- GET(
  url = "https://api.covid19api.com/dayone/country/mexico/status/confirmed")

DayOneDataText <- content(DayOneDataSa, "text")
DayOneDataTextMex <- content(DayOneDataMex, "text")

DayOneDataJson <- fromJSON(DayOneDataText, flatten = TRUE)
DayOneDataJsonMex <- fromJSON(DayOneDataTextMex, flatten = TRUE)

DayOneDfSa <- as_tibble(DayOneDataJson)
DayOneDfMex <- as_tibble(DayOneDataJsonMex)

```
We will combine the data sets for the analysis.
```{r, echo = TRUE}
SaMex <- function(df1, df2) {rbind(df1, df2)}
Day1 <- SaMex(df1 = DayOneDfSa, df2 = DayOneDfMex)

Day1$Province <- NULL
Day1$City <- NULL
Day1$CityCode <- NULL
```

These functions will allow the user to query the Day One Mexico and Day One South Africa APIs based on the columns by entering all or a column name.  To subset by rows the user can enter arguments for the filter function.
```{r, echo = TRUE, message = FALSE, results = 'hide'}
DayOneSa <- function(type = "all"){
  OutputDat <- GET("https://api.covid19api.com/dayone/country/south-africa/status/confirmed")
    DataSa <- fromJSON(rawToChar(OutputDat$content))
  if(type!="all"){DataSa <- DataSa %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date) %>% filter(Cases == 1)}
  return(DataSa)}
DayOneSa("CountryCode")
```
```{r, echo = TRUE, message = FALSE, results = 'hide'}
DayOneMex <- function(type = "all"){
  OutputDatMex <- GET("https://api.covid19api.com/dayone/country/mexico/status/confirmed")
    DataMex <- fromJSON(rawToChar(OutputDatMex$content))
  if(type!="all"){DataMex <- DataMex %>% select(type, Country, CountryCode, Cases, Lat, Lon, Date) %>% filter(Cases == 4)}
  return(DataMex)}
DayOneMex("CountryCode")
```

We will now look at summary data which contains a summary of new and total cases per country which is updated daily utilizing the steps outlined above to retrieve and parse data from the API.
```{r, echo = TRUE, message = FALSE}
resp2 <- GET("https://api.covid19api.com/summary")

resp2Text <- content(resp2, "text")
resp2Json <- fromJSON(resp2Text, flatten = TRUE)

# Save data frame as resp1Df and remove the ID column.
resp2Df <- as.data.frame(resp2Json$Countries)

resp2Df$ID <- NULL
resp2Df
```

This function allows the user to return NewDeaths, TotalDeaths,& NewRecovered, or all of the data based on inputs from the Summary API.  To subset by rows the user can enter arguments in the filter function.
```{r, echo = TRUE, message = FALSE, results = 'hide'}
Summary <- function(type = "all"){
  OutputAPI <- GET("https://api.covid19api.com/summary")
    data <- fromJSON(rawToChar(OutputAPI$content))
    data2 <- data$Countries
  if(type!="all"){data2 <- data2 %>% select(type, Country, CountryCode,  NewDeaths, TotalDeaths, NewRecovered) %>% filter(Country == "Afghanistan")}
  return(data2)}
Summary("Country")
```

The next set of functions will return live cases by case type for a country after a given date. This particular endpoint returns cases for Greece.
```{r, echo = TRUE, results = 'hide'}
Live <- GET("https://api.covid19api.com/live/country/Greece/status/confirmed/date/2020-03-21T13:13:30Z")
LiveText <- content(Live, "text")
LiveTextJson <- fromJSON(LiveText, flatten = TRUE)

# Save the data frame as LiveDf.
LiveDf <- as.data.frame(LiveTextJson)

#Remove the ID, Province, city, & CityCode columns.
LiveDf$ID <- NULL
LiveDf$Province <- NULL
LiveDf$City <- NULL
LiveDf$CityCode <- NULL
```
The following function will allow the user to query live cases for Greece by columns  To subset by rows the user can enter arguments in the filter function.
```{r, echo = TRUE, results = 'hide'}
Live2 <- function(type = "all"){
  OutputLive2 <- GET("https://api.covid19api.com/live/country/Greece/status/confirmed/date/2020-03-21T13:13:30Z")
    DataLive <- fromJSON(rawToChar(OutputLive2$content))
  if(type!="all"){DataLive <- DataLive %>% select(type, Country, CountryCode, Lat, Lon, Confirmed, Deaths, Recovered, Active, Date) %>% filter(Deaths == "12604")}
  return(DataLive)}
Live2("Country")
```

The next set of functions will access the Day One endpoint which has information on all cases by case type for a country from the first recorded case.  This function accesses cases for Denmark.  
```{r, echo = TRUE, results = 'hide', message = FALSE}
One <- GET("https://api.covid19api.com/dayone/country/denmark/status/confirmed")
OneText <- content(One, "text")
OneTextJson <- fromJSON(OneText, flatten = TRUE)

# Save the data frame as OneDf.
OneDf <- as.data.frame(OneTextJson)
OneDf
```

The following function will allow the user to query day one cases for Denmark by columns.  To subset the rows the user can enter arguments in the filter function.
```{r, echo = TRUE, results = 'hide'}
OneFun <- function(type = "all"){
  OutputOneFun <- GET("https://api.covid19api.com/dayone/country/denmark/status/confirmed")
    DataOneFun <- fromJSON(rawToChar(OutputOneFun$content))
  if(type!="all"){DataOneFun <- DataOneFun %>% select(type, Country, CountryCode, Lat, Lon, Date, Status, Cases, Province, City, CityCode) %>% filter(Province == "Faroe Islands")}
  return(DataOneFun)}
OneFun("Country")
```

The last function will contact the Countries API which returns all the available countries and provinces listed on the API which may be useful for the user when querying the returned data. 
```{r, echo = TRUE, message = FALSE}
Ctry <- GET("https://api.covid19api.com/countries")
CtryText <- content(Ctry, "text")
CtryTextJson <- fromJSON(CtryText, flatten = TRUE)

# Save the data frame as CtryDf.
CtryDf <- as.data.frame(CtryTextJson)
CtryDf
```
The following function will allow the user to query the Countries API by column.  To subset by rows the user can enter arguments in the filter function.
```{r, echo = TRUE, results = 'hide'}
CtryFun <- function(type = "all"){
  OutputCtryFun <- GET("https://api.covid19api.com/countries")
    CtryDataFun <- fromJSON(rawToChar(OutputCtryFun$content))
  if(type!="all"){CtryDataFun <- CtryDataFun %>% select(type, Country, Slug, ISO2) %>% filter(Country == "Poland")}
  return(CtryDataFun)}
CtryFun("Country")
```


# Data Analysis
We will create a contingency table that shows the occurrences of confirmed cases  between March - April 2020 for Switzerland and Norway.  Each country had a confirmed case during this time frame.
```{r, echo = TRUE}
tbl <- table(Combo$Country, Combo$Status)
tbl
```

Below is a box plot representing confirmed cases in Norway and Switzerland between March - April 2020.  The output shows that Switzerland had a higher number of confirmed cases.
```{r, echo = TRUE}
ggplot(Combo, aes(x = Cases, y = Country)) +
geom_boxplot() + geom_jitter(aes(color = Status)) + ggtitle("Boxplot for Confirmed Cases")

```

The following code calculates numerical summaries for daily cases confirmed for the two countries.  The mean case count per day for Norway was 1,759 cases and the mean case count per day for Switzerland was 5,402 cases. 
```{r, echo = TRUE}
Combo %>%
  group_by(Country) %>%
  summarize(Avg = mean(Cases), Sd = sd(Cases), Median = median(Cases), IQR =    IQR(Cases))
```
The following bar plot reports the number of days with confirmed cases for Norway & Switzerland which were similar through March - April 2020.
```{r, echo = TRUE}
ggplot(Combo, aes(x = Country)) + geom_bar(aes(fill = Status), position = "dodge") + xlab("Country") + scale_fill_discrete(name = "") + ggtitle("Confirmed Case Statuses")
```



The following contingency table reports the number of confirmed case statuses  for South Africa and Mexico. 
```{r, echo = TRUE}
tbl2 <- table(Day1$Status, Day1$Country)
tbl2
```

The following bar graph reports the number of confirmed case statuses for South Africa and Mexico since day one which were similar.  
```{r, echo = TRUE}
ggplot(Day1, aes(x = Country)) + geom_bar(aes(fill = Status), position = "dodge") + xlab("Country") + scale_fill_discrete(name = "") + ggtitle("Confirmed Case Statuses")

```

The following code calculates numerical summaries for daily cases confirmed since Day 1 for the two countries. These numbers are updated frequently.
```{r, echo = TRUE}
Day1 %>%
  group_by(Country) %>%
  summarize(Avg = mean(Cases), Sd = sd(Cases), Median = median(Cases), IQR =    IQR(Cases))
```

The following code creates a new variable (NewRatio = TotalDeaths/Total Confirmed) that represents the ratio of total deaths to total confirmed cases and appends to the Resp2Df which is based on total cases for Countries that is updated daily.
```{r, echo = TRUE}
resp2Df <- resp2Df %>% mutate(Ratio = TotalDeaths/TotalConfirmed)
head(resp2Df)
```

The following code returns numerical summaries for total deaths among all countries reported in the Summary API.  The average total deaths was 24,941.
```{r, echo = TRUE}
resp2Df %>%
    summarize(Avg = mean(TotalDeaths), Max = max(TotalDeaths), Sd = sd(TotalDeaths), Min = min(TotalDeaths))

```

The following code returns a scatter plot for the summary data set.  The plot shows there is minimal data on Newly Confirmed Cases and New Deaths; however, the graph shows a positive correlation between newly confirmed cases and new deaths.
```{r, echo = TRUE}
correlation <- cor(resp2Df$NewConfirmed, resp2Df$NewDeaths)
g <- ggplot(resp2Df, aes(x = NewConfirmed, y = NewDeaths))+ labs(y="New Deaths", x = "New Confirmed")
g + geom_point(col = "Red") + ggtitle("New Confirmed Cases vs New Deaths") + geom_text(x = 30000, y = 200, size = 5, label = paste0("Correlation = ", round(correlation, 2)))
```

The following code creates a new variable which calculates the confirmed cases to recovered ratio for Greece.  The confirmed to recovered ratios were approximately 5%. 
```{r, echo = TRUE}
LiveDf <- LiveDf %>% mutate(Ratio = Confirmed/Recovered)
head(LiveDf)
```

The following code returns numerical summaries for active cases in Greece.  Average active cases total 478,105.
```{r, echo = TRUE}
LiveDf %>%
  summarize(Avg = mean(Active), Sd = sd(Active), Median = median(Active), IQR = IQR(Active))
```

The following code will produce a histogram that displays deaths in Greece. Daily reports have remained around a cumulative total of 12K to 13K.
```{r, echo = TRUE}
ggplot(data = LiveDf, aes (Deaths)) + geom_histogram(breaks = seq(12500, 14900, by = 300), col = "blue", fill = "purple", alpha = .2) + labs(title = "Histogram for Deaths")
```

The following code returns a contingency table that reports the number of confirmed case statuses by province for the country of Denmark.  There were 578 confirmed cases in Faro Islands and 566 in Greenland.
```{r, echo = TRUE}
tbl3 <- table(OneDf$Status, OneDf$Province)
tbl3
```

# Conclusion
We have reached the conclusion of this vignette.  In summary this vignette details the functions used and created to access and parse data from the Covid19 API.  The vignette also includes an exploratory data analysis which includes plots, tables, and numerical summaries.  This vignette can be used to get a basic understanding of how to contact and retrieve data from an API.